Software Architectures for Integrating Workflow Engines in Science Gateways

### 2016-05-13 paper.tex
up to 2.5

@@

\begin{figure*}
\centering
\hspace*{0.2\columnwidth}
\subfloat[Tight integration]{
\def\svgwidth{0.6\columnwidth}
\input{figures/tight.pdf_tex}
\label{archi:tight}
}
\hfill
\subfloat[Service invocation]{
\def\svgwidth{0.6\columnwidth}
\input{figures/service.pdf_tex}
\label{archi:service}
\hspace*{0.2\columnwidth}
}\\
\hspace*{0.2\columnwidth}
\subfloat[Task encapsulation]{
\def\svgwidth{0.6\columnwidth}
\input{figures/sub-task.pdf_tex}
\label{archi:sub-task}
}
\hfill
\subfloat[Pool]{
\def\svgwidth{0.6\columnwidth}
\input{figures/agent.pdf_tex}
\label{archi:agent}
\hspace*{0.2\columnwidth}
}\\
\subfloat[Nested workflows. Left: abstract model. Right: instantiation with service invocation.]{
\def\svgwidth{0.9\columnwidth}
\input{figures/nested-1.pdf_tex}
\def\svgwidth{0.9\columnwidth}
\input{figures/nested-2.pdf_tex}
\label{archi:nested}
}\\
\subfloat[Workflow import. Left: abstract model. Right: instantiation with service invocation.]{
\def\svgwidth{0.5\columnwidth}
\input{figures/import-1.pdf_tex}
\hfill
\def\svgwidth{0.5\columnwidth}
\input{figures/import-2.pdf_tex}
\label{archi:import}
}
\caption{Architectures}
\label{fig:architectures}
\end{figure*}
\begin{table*}
\centering
\footnotesize
\rowcolors{1}{white}{gray!25}
\begin{tabular}{ll}
  \textbf{Architecture} & \textbf{Systems} \\
  \hline
  Tight & \pbox{1.5\columnwidth}{
          Catania Science Gateway Framework~\cite{ardizzone2012decide},
          Distributed application runtime environment (DARE~\cite{maddineni2012distributed}),
          DECIDE~\cite{ardizzone2012decide}$^{c,n}$, LONI Pipeline Environment$^n$~\cite{dinov2009efficient}
          }
  \\
  Service & \pbox{1.5\columnwidth}{
            Apache Airvata~\cite{marru2011apache}, e-BioInfra~\cite{shahand2015data}$^{w,n}$, HubZero with Pegasus~\cite{CPE:CPE3257}, MoSGrid~\cite{kruger2014mosgrid}$^w$, System in~\cite{wu2010accelerating}, Vine Toolkit~\cite{DBLP:journals/scpe/SzejnfeldDKKKKLPTWDNW10}, Virtual Imaging Platform~\cite{GLAT-13}$^n$, WS-PGRADE/gUSE framework~\cite{Kacsuk2012},
            Science gateways in~\cite{kacsuk2014science}$^w$
            } \\
  Task & CBRAIN and PSOM~\cite{GLAT-16}$^n$, CBRAIN and FSL$^n$\\
  Pool & SHIWA pool~\cite{ROGE-13}\\
  Nested & \pbox{1.5\columnwidth}{SHIWA Simulation Platform (Coarse-Grained Interoperability~\cite{terstyanszky2014enabling})$^w$, HubZero with Pegasus (via hierarchical workflows)~\cite{Deelman201517}, Tavaxy~\cite{Abouelhoda2012}.
           }\\
  Import & SHIWA Simulation Platform (Fine-Grained Interoperability)~\cite{plankensteiner-prodan-etal:2013}$^w$, Tavaxy~\cite{Abouelhoda2012}, system in~\cite{delaGarza2016}.
\end{tabular}
\caption{Classification of science gateways based on the architecture used to integrate workflow engines. $^c$: based on the Catania Science Gateway Framework. $^w$: based on WS-PGRADE/gUSE. $^n$: used for neuroimaging.}
\label{table:system-classification}
\end{table*}

@@

See Figure~\subref*{archi:service}. The workflow engine is available
externally to the science gateway, as a service. The science gateway
controls the service through a specific interaction (\texttt{d}) that
might be implemented as a web-service call (e.g., RESTful or SOAP), as
a command-line or as any other method that offers a well-defined
interface to the workflow engine. The workflow engine might be invoked
either as a black box that completely masks the infrastructure and
tasks, or as a white box that allows for some interaction with
them. The workflow engine is responsible for controlling the tasks on
the infrastructure (\texttt{b}) and for performing the required data
transfers to execute them (\texttt{c$_2$}). User data is usually
managed through the science gateway (\texttt{c$_1$}), although it
might as well be delivered by the workflow engine directly to the
user.

This architecture is largely adopted, in systems such as Apache
Airvata~\cite{marru2011apache}, Vine
Toolkit~\cite{DBLP:journals/scpe/SzejnfeldDKKKKLPTWDNW10}, Virtual
Imaging Platform~\cite{GLAT-13}, the system
in~\cite{wu2010accelerating} and the WS-PGRADE/gUSE
framework~\cite{Kacsuk2012}. The integration between the
HubZero science gateway and the Pegasus workflow engine performed
in~\cite{CPE:CPE3257} also uses a service architecture where the
\texttt{d} interaction is implemented as a set of calls to Pegasus'
command-line tools (e.g., pegasus-status, pegasus-plan,
etc.). Figure~\ref{fig:vip-architecture} shows the architecture of the
Virtual Imaging Platform, where the MOTEUR workflow
engine~\cite{GLAT-08i} is invoked through a service.

@@

The science gateway also transfers both user and workflow data across
the infrastructure, so that interactions \texttt{c$_1$} and
\texttt{c$_2$} are both covered by \texttt{c}. In practice, both
\texttt{c$_1$} and \texttt{c$_2$} can be implemented using the same
pieces of code.

Task encapsulation is implemented in CBRAIN~\cite{SHER-14} where it is
used to integrate the FSL toolkit~\cite{Jenkinson2012782} and the PSOM
workflow engine~\cite{bellec2012pipeline}. The CBRAIN-FSL integration
allows to leverage FSL pipelines written in low-level workflow
languages (Linux executables and scripts) that submit tasks uniformly
through a specific tool called fsl\_sub. It is diagrammed in
Figure~\subref*{fig:cbrain-fsl-architecture}, where the science gateway is
represented by components \texttt{CBRAIN portal} and \texttt{CBRAIN
  execution server} and the infrastructure consists of \texttt{Storage
  servers} and \texttt{Computing cluster with shared file system}.

>>>>

\note{Sil}{i changed above a few occurrences of "diagrammed". i know it is correct, but i am not used to seeing it. this is a matter of style, just forget my changes below then}

@@

\begin{figure*}
\centering
\hspace*{0.2\columnwidth}
\subfloat[Tight integration]{
\def\svgwidth{0.6\columnwidth}
\input{figures/tight.pdf_tex}
\label{archi:tight}
}
\hfill
\subfloat[Service invocation]{
\def\svgwidth{0.6\columnwidth}
\input{figures/service.pdf_tex}
\label{archi:service}
\hspace*{0.2\columnwidth}
}\\
\hspace*{0.2\columnwidth}
\subfloat[Task encapsulation]{
\def\svgwidth{0.6\columnwidth}
\input{figures/sub-task.pdf_tex}
\label{archi:sub-task}
}
\hfill
\subfloat[Pool]{
\def\svgwidth{0.6\columnwidth}
\input{figures/agent.pdf_tex}
\label{archi:agent}
\hspace*{0.2\columnwidth}
}\\
\subfloat[Nested workflows. Left: abstract model. Right: instantiation with service invocation.]{
\def\svgwidth{0.9\columnwidth}
\input{figures/nested-1.pdf_tex}
\def\svgwidth{0.9\columnwidth}
\input{figures/nested-2.pdf_tex}
\label{archi:nested}
}\\
\subfloat[Workflow import. Left: abstract model. Right: instantiation with service invocation.]{
\def\svgwidth{0.5\columnwidth}
\input{figures/import-1.pdf_tex}
\hfill
\def\svgwidth{0.5\columnwidth}
\input{figures/import-2.pdf_tex}
\label{archi:import}
}
\caption{Architectures}
\label{fig:architectures}
\end{figure*}
\begin{table*}
\centering
\footnotesize
\rowcolors{1}{white}{gray!25}
\begin{tabular}{ll}
  \textbf{Architecture} & \textbf{Systems} \\
  \hline
  Tight & \pbox{1.5\columnwidth}{
          Catania Science Gateway Framework~\cite{ardizzone2012decide},
          Distributed application runtime environment (DARE~\cite{maddineni2012distributed}),
          DECIDE~\cite{ardizzone2012decide}$^{c,n}$, LONI Pipeline Environment$^n$~\cite{dinov2009efficient}
          }
  \\
  Service & \pbox{1.5\columnwidth}{
            Apache Airvata~\cite{marru2011apache}, Neurosciece Gateway~\cite{shahand2015data}$^{w,n}$, HubZero with Pegasus~\cite{CPE:CPE3257}, MoSGrid~\cite{kruger2014mosgrid}$^w$, System in~\cite{wu2010accelerating}, Vine Toolkit~\cite{DBLP:journals/scpe/SzejnfeldDKKKKLPTWDNW10}, Virtual Imaging Platform~\cite{GLAT-13}$^n$, WS-PGRADE/gUSE framework~\cite{Kacsuk2012},
            Science gateways in~\cite{kacsuk2014science}$^w$
            } \\
  Task & CBRAIN and PSOM~\cite{GLAT-16}$^n$, CBRAIN and FSL$^n$\\
  Pool & SHIWA pool~\cite{ROGE-13}\\
  Nested & \pbox{1.5\columnwidth}{SHIWA Simulation Platform (Coarse-Grained Interoperability~\cite{terstyanszky2014enabling})$^w$, HubZero with Pegasus (via hierarchical workflows)~\cite{Deelman201517}, Tavaxy~\cite{Abouelhoda2012}.
           }\\
  Import & SHIWA Simulation Platform (Fine-Grained Interoperability)~\cite{plankensteiner-prodan-etal:2013}$^w$, Tavaxy~\cite{Abouelhoda2012}, system in~\cite{delaGarza2016}.
\end{tabular}
\caption{Classification of science gateways based on the architecture used to integrate workflow engines. $^c$: based on the Catania Science Gateway Framework. $^w$: based on WS-PGRADE/gUSE. $^n$: used for neuroimaging.}
\label{table:system-classification}
\end{table*}

@@

See Figure~\subref*{archi:service}. The workflow engine is available
externally to the science gateway, as a service. The science gateway
controls the service through a specific interaction (\texttt{d}) that
might be implemented as a web-service call (e.g., RESTful or SOAP), as
a command-line or as any other method that offers a well-defined
interface to the workflow engine. The workflow engine might be invoked
either as a black box that completely masks the infrastructure and
tasks, or as a white box that allows for some interaction with
them. The workflow engine is responsible for controlling the tasks on
the infrastructure (\texttt{b}) and for performing the required data
transfers to execute them (\texttt{c$_2$}). User data is usually
managed through the science gateway (\texttt{c$_1$}), although it
might as well be delivered by the workflow engine directly to the
user.
\note{Sil}{about c1: should it also show from engine to infra in the figure? this is what i understand from the above sentence}

This architecture is largely adopted in systems such as Apache
Airvata~\cite{marru2011apache}, Vine
Toolkit~\cite{DBLP:journals/scpe/SzejnfeldDKKKKLPTWDNW10}, Virtual
Imaging Platform~\cite{GLAT-13}, the systems
in~\cite{wu2010accelerating,shahand:2012jgc} and the WS-PGRADE/gUSE
framework~\cite{Kacsuk2012}. The integration between the
HubZero science gateway and the Pegasus workflow engine performed
in~\cite{CPE:CPE3257} also uses a service architecture, where the
\texttt{d} interaction is implemented as a set of calls to Pegasus'
command-line tools (e.g., pegasus-status, pegasus-plan,
etc.). Figure~\ref{fig:vip-architecture} shows the architecture of the
Virtual Imaging Platform, where the MOTEUR workflow
engine~\cite{GLAT-08i} is invoked through a service.

@@

The science gateway also transfers both user and workflow data across
the infrastructure, so that interactions \texttt{c$_1$} and
\texttt{c$_2$} are both covered by \texttt{c}. 
\note{Sil}{is this next sentence relevant?}In practice, both
\texttt{c$_1$} and \texttt{c$_2$} can be implemented using the same
pieces of code.

Task encapsulation is implemented in CBRAIN~\cite{SHER-14} where it is
used to integrate the FSL toolkit~\cite{Jenkinson2012782} and the PSOM
workflow engine~\cite{bellec2012pipeline}. The CBRAIN-FSL integration
allows leveraging FSL pipelines written in low-level workflow
languages (Linux executables and scripts) that submit tasks uniformly
through a specific tool called fsl\_sub. This integration is shown in
Figure~\subref*{fig:cbrain-fsl-architecture}, where the science gateway is
represented by components \texttt{CBRAIN portal} and \texttt{CBRAIN
  execution server} and the infrastructure consists of \texttt{Storage
  servers} and \texttt{Computing cluster with shared file system}.
